---
title: "Intro to use getWebServiceData function"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro to use getWebServiceData function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cefasSpatialHub)
```

## Basic use of the getWebServiceData function

The getWebServiceData function retrieve data from a provided ESRI REST API URL. You need to pass a few essential parameters: 

  - web_service: The url of the REST API
  - layer_index: what layer from the API you want to query
  
This information is the essential because is pointing to where API of the data is located. 

The additional optional parameters will define your final API url and the result of it . 

```{r }
## Example of a protected url api 
url = "https://giserver.cefas.co.uk/devserver/rest/services/ThirdPartyData/MPA_UK_offshore_JNCC/MapServer/"

## Example of a open layer api and the index layer 

url_l = "https://giserver.cefas.co.uk/devserver/rest/services/fishing_activity/Landing_composition_statistics_by_species_ICES_rectangles_and_ports_2009_to_2021/MapServer/"
ly = 8 

```

### Get your data as a spatial SF object or as a data.frame

To choose if the data is obtained as a SF spatial object or a data frame has to be defined in the **spatial_layer** paramter. 
By default it is TRUE and will return the data as a SF spatial object , change to FALSE if only the data frame is required

```{r} spatial_layer = TRUE ```

### Filter your data 

The data fetched can be filtered using a SQL clause in **where_clause** parameter with the following format: 

```sql ft_lyear in ( 2019, 2020 ) ```

Otherwise if not WHERE parameter is provided will get all the records in the dataset ( which could delay the response)


```{r} 

where_query =  "ft_lyear in ( 2019, 2020 ) AND le_spe in  ( 'COD', 'HAD', 'WHG' ) "

```


### Get defined fields

Also you can select which fields you want to as response. The default value will get all the fields in the database, otherwise the lsit has to be specified in **output_fields** parameter.

```{r }

output_fields = 'english_name'

```

### Limit the number of records 

To give a first view into the data you can limit the numebr of records fetched. Otherwise the default value will retrieve all records

```{r}  n_records = 10 ```

### Get only the API string to open in other software (e.g QGIS )

You can define **return_api** to TRUE to not return any data but instead the API url . You can copy it an use it to retrieve teh data in other sofwtare. 

For example in QGIS . see example 5



## EXAMPLES OF 2 USES OF THE FUNCTION


### 1. Get the 15 first records as a SPATIAL LAYER

```{r}

  landings_map =     getWebServiceData ( web_service=   url_l,
                                         layer_index = ly ,
                                         output_fields = "*",
                                         n_records = 15,
                                         spatial_layer = T 
  )
  

```


### 2. Get all records as a DATA.FRAME

```{r}

    getWebServiceData ( web_service=   url_l,
                                         layer_index = ly ,
                                         output_fields = "*",
                                         spatial_layer = F 
  )
  

```

### 3. Fitler the data by SQL clause and return data as SPATIAL LAYER 
```{r}

    getWebServiceData ( web_service=   url_l,
                                         layer_index = ly ,
                                         where_clause = where_query,
                                         output_fields = "*",
                                         spatial_layer = T 
  )
  

```


### 4. Get unique values of a field  
```{r}

      getWebServiceData ( web_service=   url_l,
                                         layer_index = ly ,
                                         output_fields = "english_name",
                                         dist_values = T,
                                         spatial_layer = F 
  )
  

```


### 5. Get only the constructed API string to open somewhere else (e.g. QGIS)
```{r}

      getWebServiceData ( web_service=   url_l,
                                         layer_index = ly ,
                                         output_fields = "*",
                                         n_records = 5,
                                         spatial_layer = T , 
                                         return_api = T
  )
  

```
